Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-09 p2/axiom-proof.tex
Simplified the proofs of rules

In this section we derive the \namecrefs{ax_separate_commute} listed in
\cref{rules_lemma} from our filesystem model.
For the purposes of the proofs we extend the notion of filesystems 
to the full set of functions mapping $\setn$ to $\setv$, even if they
do not have the tree property.
This is so that we can describe the values they would acquire under certain nodes,
and pinpoint where they may be violating the tree property.
If they do, we continue to say that the filesystem is broken,
and say that the filesystem is \emph{broken at $n$} if 
$n$ is not a dictionary, but has at least one non-empty child.
We continue to require that filesystems are not empty at finitely many nodes.

Accordingly, where needed, we also implicitly extend filesystem commands
so that they would return a filesystem even if it violates the tree property.
We continue to say that they break the filesystem in these cases,
and commands continue to be undefined where their input types are incompatible
with the input filesystem.
We note that once a filesystem is broken, it stays broken, even if
after a subsequent command it would no longer violate the tree property.

It follows from the definition of the tree property that
whether a filesystem is \emph{broken at a node $n$} is determined only
by the values at $n$ and the children of $n$.
In other words,

\begin{myclm}[Locality of tree property violations]\mlabel{broken_local}
If two filesystems over $\setn$ have the same values at a node $n$,
and at all children of $n$ (if they exist),
then either both filesystems are broken at $n$, or none of them are.
\end{myclm}

@@

In the former case, based on \cref{broken_local}, 
$\czwmv\aFS$ must be broken at $m$ or $\parentf{m}$, since it only changed at $m$.
Let us look at the first case.
Since $n\indep m$, we know that the parent and children of $m$ have the same value
in $\FS$ as they do in $\cxynv\aFS$. However, this is a contradiction,
as applying $\czwmv$ to $\FS$ leads to a broken filesystem,
but applying it to $\cxynv\aFS$ (from the left side) does not.

@@

\begin{proof}
Suppose $\cxynnv$ and $\czwnv$ satisfy the conditions,
and there is an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czwnv\neq\cbba{n}$, then either $\FS(n)\neq\empty$ or $\czwnv\aFS(n)\neq\empty$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxynnv\neq\cdda{\nn}$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satisfy the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.
\end{proof}

@@

\begin{proof}
As for an assertion command $\cxynv$, $x=y$, we see that it either
breaks a filesystem, or leaves it in the same state.
\end{proof}

>>>>

In this section we derive the \namecrefs{ax_separate_commute} listed in
\cref{rules_lemma} from our filesystem model.
In the proofs, we may say that a filesystem
is \emph{broken at $n$}
if after applying a command, it violates the tree property
because
$n$ is not a dictionary, but has at least one non-empty child.

% For the purposes of the proofs we extend the notion of filesystems
% to the full set of functions mapping $\setn$ to $\setv$, even if they
% do not have the tree property.
% This is so that we can describe the values they would acquire under certain nodes,
% and pinpoint where they may be violating the tree property.
% If they do, we continue to say that the filesystem is broken,
% and say that the filesystem  if 
% We continue to require that filesystems are not empty at finitely many nodes.
% 
% Accordingly, where needed, we also implicitly extend filesystem commands
% so that they would return a filesystem even if it violates the tree property.
% We continue to say that they break the filesystem in these cases,
% and commands continue to be undefined where their input types are incompatible
% with the input filesystem.
% We note that once a filesystem is broken, it stays broken, even if
% after a subsequent command it would no longer violate the tree property.

It follows from the definition of the tree property that
whether a filesystem is \emph{broken at a node $n$} is determined only
by the values at $n$ and the children of $n$.
% In other words,
% 
% \begin{myclm}[Locality of tree property violations]\mlabel{broken_local}
% If two filesystems over $\setn$ have the same values at a node $n$,
% and at all children of $n$ (if they exist),
% then either both filesystems are broken at $n$, or none of them are.
% \end{myclm}

@@

In the former case 
$\czwmv\aFS$ must be broken at $m$ or $\parentf{m}$, since it only changed at $m$.
Let us look at the first case.
Since $n\indep m$, we know that the parent and children of $m$ have the same value
in $\FS$ as they do in $\cxynv\aFS$. However, this is a contradiction,
as applying $\czwmv$ to $\FS$ leads to a broken filesystem,
but applying it to $\cxynv\aFS$ (from the left side) does not.

@@

\begin{proof}
We use an inverse proof and
suppose $\cxynnv$ and $\czwnv$ satisfy the conditions,
and there is an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czwnv\neq\cbba{n}$, then either $\FS(n)\neq\empt$ or $\czwnv\aFS(n)\neq\empt$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxynnv\neq\cdda{\nn}$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satisfy the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.
\end{proof}

@@

\begin{proof}
The two assertion commands $\cdda{n}$ and $\cbba{n}$ either
breaks a filesystem, or leaves it in the same state.
\end{proof}
